### 2016-04-12 Chapters/chap1/introduction.tex
Gonna give it to the committee

Energy and electronic transport plays a central role in a
wide range of  chemical  and biological systems.
It is the fundamental mechanism for transporting the  energy  of a
absorbed photon to a reaction center in light harvesting systems
and for initiating a wide range of photo-induced  chemical processes,
including vision, DNA mutation, and pigmentation.
The seminal model for calculating electron transfer rates was developed by
Marcus in the 1950's\cite{marcus1956theory,marcus1965theory,marcus1993electron}.
\begin{equation}
k_{Marcus}=\frac{2\pi}{\hbar}|V_{ab}|^{2}\frac{1}{\sqrt{4\text{\ensuremath{\pi}}k_{B}T\lambda}}e^{-(\lambda+ \Delta\epsilon)^{2}/4\text{\ensuremath{\lambda}}k_{B}T}.\label{eq:marcus}
\end{equation}
where $\lambda$ is  energy required to reorganize the environment
following the transfer of an electron from donor to acceptor.
and $\Delta \epsilon$ is the driving force for the reaction.

@@

If we assume that the nuclear motions about the equilibrium configurations of the
donor and acceptor species is harmonic,  the chemical reactions resulting from
energy or charge transfer events can be understood in terms of intersecting
diabatic potentials as sketched in Fig.~\ref{marcus}.    The upper and lower
curves are the adiabatic potential energy surfaces describing the nuclear dynamics
resulting from an energy or charge transfer event, taking the geometry of the donor
state as the origin.
%For reactions carried out in open systems,  the Marcus parabolas represent
%free energy surfaces with driving force $\Delta G$.
As the transfer occurs by crossing an energy barrier,
the transfer rate can be expected to be in the Arrhenius form
\begin{eqnarray}
k\propto e^{-E_{A}/k_{B}T},
\end{eqnarray}
with $E_{A}$ as the activation energy.
 Using $E_{A}={(\lambda+\Delta \epsilon)^{2}}/{4\lambda}$
 we can relate the activation energy to both the reorganization energy and driving force, $-\Delta \epsilon$.
One of the most profound predictions of the theory is that  as the driving force increases,
the transfer rate reaches a maximum and further increases in the driving force
lead to lower reaction rates, termed the inverted regime.

A number of years ago,  we developed a time-convolutionless  master equation approach for computing
state-to-state rates in which the coupling between  states depends upon the
nuclear coordinates\cite{pereverzev2006time}. This approach incorporates a fully quantum
mechanical treatment of both the nuclear and electronic degrees of freedom and recovers
well-known Marcus expression in the semiclassical limit.  The model is parameterized by the
vibrational normal mode frequencies, and the electronic energies and energy derivatives
at a reference configuration.  The approach has been used by our group to compute state-to-state
transition rates in semi-empirical models for organic semiconducting light-emitting diode and photovoltaics
\cite{tamura2008phonon,tamura2007exciton,bittner2014noise,singh2009fluorescence}.
This paper represents the first time we have used this approach within the context of a fully {\em ab initio} quantum
chemical model.   As such, this present work provides an important benchmark of the approach since all parameters will be determined
using state-of-the-art quantum chemical methods
and results compared to both theoretical and experimental
rates.

@@

We now separate Eq.~\ref{ham1} into diagonal and
off-diagonal terms
\begin{eqnarray}
\hat  H = \hat H_{o} + \hat V
\end{eqnarray}
and perform a polaron transform
using the unitary transformation~\cite{grover1970exciton,rice1994excitons,pereverzev2006time}.
\begin{eqnarray}
U&=&e^{-\sum_{ni}\!\!\frac{g_{nni}}{\hbar\omega_i}|n\rangle \langle
n|(a^{\dagger}_i-a_i)}
 \nonumber \\
&=&
\sum_{n}|n\rangle \langle n|e^{-\sum_{i}\!\!\frac{g_{nni}}{\hbar\omega_i}(a^{\dagger}_i-a_i)}
\label{unitary}
\end{eqnarray}
under which the transformed Hamiltonian is written in terms of the
diagonal elements
\begin{eqnarray} \tilde H_0=U^{-1}H_0U
=\sum_n\tilde\epsilon_n |n\rangle \langle
n|+\sum_i\omega_ia^{\dagger}_ia_i,
 \end{eqnarray}
with  the renormalized electronic energies,
\begin{eqnarray}
\tilde\epsilon_n=\epsilon_n-\sum_{i}\frac{g_{nni}^2}{\hbar\omega_i},
\end{eqnarray}
and off-diagonal terms,
\begin{eqnarray} \hat V_{nm}=\sum_{i}g_{nmi}\left(a^{\dagger}_i+
a_i-\frac{2g_{nni}}{\hbar\omega_i}\right)e^{\sum_{j}\frac{(g_{nnj}-g_{mmj})}{\hbar\omega_j}(a^{\dagger}_j-a_j)}.
\label{opm}
\end{eqnarray}
In the transformed (or dressed) picture the electronic transition from state
$|n\rangle$ to $|m\rangle$ is accompanied by the excitations of all the
normal phonon modes.  Transforming to the interaction representation
and performing a trace over the phonons gives the spectral density in
terms of the autocorrelation of the electron-phonon coupling
operators.
\begin{eqnarray}
S_{nm}(\tilde\omega) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle.\label{spec-dens}
\end{eqnarray}
Here, $\hat V_{nm}(t)$ is the electron-phonon coupling term in the Heisenberg representation and
$\langle \cdots \rangle$ denotes a thermal average over the
vibrational degrees of freedom.
The derivation and explicit form for the kernel in Eq.~\ref{spec-dens}  is quite lengthy and is given in
Ref.~\cite{pereverzev2006time}.

\subsection{Non-Markovian Master Equation and Golden-Rule rates}

@@

At this point it is useful to connect the various terms in the phonon-dressed Hamiltonian
with specific physical parameters.
First, the reorganization energy is given by
%The terms involving $$
%are related to the reorganized energy difference between the two states.
%$$
%\lambda_{nm} =\sum_{j}((g_{nnj}-g_{mmj})/\omega_j)^{2} = \sum_{j}\hbar \omega_{j}S_{j}
%$$
\[
\lambda_{nm}=\sum_{j}\frac{\left(g_{nnj}-g_{mmj}\right)^{2}}{\omega_{j}} = \sum_{j}\hbar \omega_{j}S_{j}
\]
where  the $\{S_{j}\}$  are  the Huang-Rhys factors for each phonon mode.
These are related to the Franck-Condon factor
describing the overlap between the $v_j=1$ vibronic state in one electronic state
with the $v_j=0$ vibronic state in the other.
Likewise, the energy difference between the renormalized energy gaps is related to the
driving force of the state-to-state transition,
\begin{eqnarray}
\Delta E_{nm} = \tilde \epsilon_n-\tilde \epsilon_m.
\end{eqnarray}

We first developed the approach about 10 years ago. However, the difficulty in using this approach is that
it requires both diagonal (${\mathbf g}_{nn}$) and off-diagonal  (${\mathbf g}_{nm}$) derivative couplings between
adiabatic states.    In the following section, we discuss how we have used the Edmiston-Ruedenberg localization scheme
to estimate the couplings\cite{edmiston1963localized}.   We also
present how one can construct a reduced set of harmonic modes that
fully capture the electron/nuclear coupling.
\subsection{Parameterization From Quantum Chemistry}

@@

\subsection{Lanczos Method}
It is crucial to notice that the vectors  given in Eq.~\ref{eq:locaiHam} are {\em not linearly independent}.
 Consequently, special care
must be taken to generate the reduced sub-space.  To do so, we use an iterative Lanczos approach
taking the normalized vector ${\mathbf v}_{1} = {\mathbf g}_{22}$ as a starting point.

As above, we initialize each step indexed by $k$,  by defining a projection operator
\begin{eqnarray}
{\mathbf P}_{k} = {\mathbf v}_{k}\otimes  {\mathbf v}_{k}
\end{eqnarray}
and its complement ${\mathbf Q}_{k} = {\mathbf I}  - {\mathbf P}_{k}$.
%${\mathbf P}_{k}$ is the projection operator for the
$k$-th mode.
%We also construct
%\begin{eqnarray}
%\mathbf p = \sum_{k} \mathbf P_{k}
%\end{eqnarray}
%as  the total projection operator for all ${k} \le N$ modes.
We then project the hessian matrix $\mathbf\Omega$ into each subspace {\em viz.}
\begin{eqnarray}
\mathbf\Omega_{p} = \mathbf P_{k}\cdot \mathbf\Omega \cdot \mathbf P_{k} \,\, \& \,\, \mathbf\Omega_{q} = \mathbf Q_{k}\cdot \mathbf\Omega \cdot  \mathbf Q_{k}
\end{eqnarray}
and diagonalize each to obtain eigenvalues and eigenvectors $\{\alpha_{p}, {\mathbf M}_{p}\}$ and $\{\alpha_{q}, {\mathbf M}_{q}\}$
respectively.
As above, $\mathbf\Omega_{p} $ and $\mathbf\Omega_{q}$ are $N\times N$ matrices.
The first set will have a single
non-trivial eigenvalue and the second set
will have $N-k$ non-trivial eigenvalues.  As above we collect the non-trivial eigenvectors associated with each
to form the orthogonal transformation matrix
\begin{eqnarray}
{\mathbf M}_{k} = \{{\mathbf M}_{p},{\mathbf M}_{q}\}.
\end{eqnarray}
and again transform the full hessian $\mathbf\Omega$ into this new vector space to form the $N\times N$ matrix $\mathbf\Omega'$.
 At each step in the iteration, the transformed hessian, $\mathbf\Omega'$ is in the form of a
$k\times k$ tri-diagonal submatrix in the upper-left part of the matrix and
a diagonal submatrix in the lower-right.  For example, after $k=3$ iterations, one has
a Hessian matrix of the form:
\begin{eqnarray}
{\mathbf\Omega}' =
\begin{pmatrix}
\alpha_{1}   & b_{1}    & 0                    &               &             &            & 0 \\
b_{1}     & \alpha_{2}  & b_{2}                         \\
0            &   b_{2}         & \alpha_{3}    &   c_{k+1}   & c_{k+2}  & \cdots & c_{N}     \\
               &                     &           c_{k+1}           &     \alpha_{k+ 1} & &             & 0                 \\
               &                     &           c_{k+2}           &                 & \alpha_{k+2}  \\
               &                     &          \vdots           &                 &                    &  \ddots \\
   0           &                     &           c_{N}           &     0   &&& \alpha_{N}\\
\end{pmatrix}.
\label{omega-prime}
\end{eqnarray}
We note that only the $k$-th mode is coupled the $N-k$ remaining modes.
Since all of the transformations are orthogonal, diagonalizing $\mathbf\Omega'$ at any point
returns the original Hessian matrix.

@@

At any point along the way, we can terminate the iteration and obtain a reduced set of
couplings.  Since the  Lanczos approach uses the power method for finding the largest eigenvector of a matrix,
it converges first upon the vector with the largest electron/nuclear coupling--which
 we refer to as the ``primary mode''.  Subsequent iterations produce reduced modes with
 weaker electron/nuclear couplings and the entire  process can be terminated after a few iterations.
After $k$-steps, the final electron-phonon couplings are then obtained by
projecting the original set of couplings (in the normal mode basis) into the final vector space.

For the first iteration, ${\mathbf v}_{1}$ is parallel to the bare electron-phonon coupling vector $g_{22}$
and the associated frequency is ${\mathbf v}_{1}\cdot\Omega\cdot{\mathbf v}_{1}$.   The subsequent iterations introduce
corrections to this via phonon-phonon coupling.  For example, for the $k=3$ iteration,
we would determine the active vector space in terms of the upper-left 3$\times 3$ block of the matrix in
Eq.~\ref{omega-prime}.
\begin{eqnarray}
{\mathbf \Omega}_{3}' =
\begin{pmatrix}
\alpha_{1}   & b_{1}    & 0      \\
b_{1}     & \alpha_{2}  & b_{2}                         \\
0            &   b_{2}         & \alpha_{3}
\end{pmatrix}
\end{eqnarray}
Diagonalizing  ${\mathbf \Omega}_{3}'$ returns a set of frequencies and  associated eigenvectors
which are then used to compute the electron-phonon couplings in this reduced active space.
After $N-1$ iterations, $\mathbf\Omega'$ is a fully tridiagonal matrix and diagonalizing this returns the original
normal mode basis.

>>>>

Energy and electronic transport plays a central role in a
wide range of  chemical  and biological systems.
It is the fundamental mechanism for transporting the  energy  of an
absorbed photon to a reaction center in light harvesting systems
and for initiating a wide range of photo-induced  chemical processes,
including vision, DNA mutation, and pigmentation.
The seminal model for calculating electron transfer rates was developed by
Marcus in the 1950's\cite{marcus1956theory,marcus1965theory,marcus1993electron}.
\begin{equation}
k_{Marcus}=\frac{2\pi}{\hbar}|V_{ab}|^{2}\frac{1}{\sqrt{4\text{\ensuremath{\pi}}k_{B}T\lambda}}e^{-(\lambda+ \Delta\epsilon)^{2}/4\text{\ensuremath{\lambda}}k_{B}T}.\label{eq:marcus}
\end{equation}
where $\lambda$ is  energy required to reorganize the environment
following the transfer of an electron from donor to acceptor.
and $\Delta \epsilon$ is the driving force for the reaction, as illustrated in Fig.~\ref{marcus}. If we assume that the nuclear motions about the equilibrium configurations of the
donor and acceptor species is harmonic,  the chemical reactions resulting from
energy or charge transfer events can be understood in terms of intersecting
diabatic potentials as sketched.    The upper and lower
curves are the adiabatic potential energy surfaces describing the nuclear dynamics
resulting from an energy or charge transfer event, taking the geometry of the donor
state as the origin.

@@

%For reactions carried out in open systems,  the Marcus parabolas represent
%free energy surfaces with driving force $\Delta G$.
As the transfer occurs by crossing an energy barrier,
the transfer rate can be expected to be in the Arrhenius form
\begin{eqnarray}
k\propto e^{-E_{A}/k_{B}T},
\end{eqnarray}
with $E_{A}$ as the activation energy.
 Using $E_{A}={(\lambda+\Delta \epsilon)^{2}}/{4\lambda}$
 we can relate the activation energy to both the reorganization energy and driving force, $-\Delta \epsilon$.
One of the most profound predictions of the theory is that  as the driving force increases,
the transfer rate reaches a maximum and further increases in the driving force
lead to lower reaction rates, termed the inverted regime.

A number of years ago,  our group developed a time-convolutionless  master equation approach for computing
state-to-state rates in which the coupling between  states depends upon the
nuclear coordinates\cite{pereverzev2006time}. This approach incorporates a fully quantum
mechanical treatment of both the nuclear and electronic degrees of freedom and recovers the
well-known Marcus expression in the semiclassical limit.  The model is parameterized by the
vibrational normal mode frequencies, and the electronic energies and energy derivatives
at a reference configuration.  The approach has been used by our group to compute state-to-state
transition rates in semi-empirical models for organic semiconducting light-emitting diode and photovoltaics
\cite{tamura2008phonon,tamura2007exciton,bittner2014noise,singh2009fluorescence}.
Our work here represents the first time this approach has been used within the context of a fully {\em ab initio} quantum
chemical model.   As such, this present work provides an important benchmark of the approach since all parameters will be determined
using state-of-the-art quantum chemical methods
and results compared to both theoretical and experimental
rates.

@@

We now separate Eq.~\ref{ham1} into diagonal and
off-diagonal terms
\begin{eqnarray}
\hat  H = \hat H_{o} + \hat V
\end{eqnarray}
and perform a polaron transform
using the unitary transformation~\cite{grover1970exciton,rice1994excitons,pereverzev2006time}.
\begin{eqnarray}
U&=&e^{-\sum_{ni}\!\!\frac{g_{nni}}{\hbar\omega_i}|n\rangle \langle
n|(a^{\dagger}_i-a_i)}
 \nonumber \\
&=&
\sum_{n}|n\rangle \langle n|e^{-\sum_{i}\!\!\frac{g_{nni}}{\hbar\omega_i}(a^{\dagger}_i-a_i)}
\label{unitary}
\end{eqnarray}
under which the transformed Hamiltonian is written in terms of the
diagonal elements
\begin{eqnarray} \tilde H_0=U^{-1}H_0U
=\sum_n\tilde\epsilon_n |n\rangle \langle
n|+\sum_i\omega_ia^{\dagger}_ia_i,
 \end{eqnarray}
with  the renormalized electronic energies,
\begin{eqnarray}
\tilde\epsilon_n=\epsilon_n-\sum_{i}\frac{g_{nni}^2}{\hbar\omega_i},
\end{eqnarray}
and off-diagonal terms,
\begin{eqnarray} \hat V_{nm}=\sum_{i}g_{nmi}\left(a^{\dagger}_i+
a_i-\frac{2g_{nni}}{\hbar\omega_i}\right)e^{\sum_{j}\frac{(g_{nnj}-g_{mmj})}{\hbar\omega_j}(a^{\dagger}_j-a_j)}.
\label{opm}
\end{eqnarray}
In the transformed (or dressed) picture the electronic transition from state
$|n\rangle$ to $|m\rangle$ is accompanied by the excitations of all the
normal modes. Following the conventional notations in this field, we refer to all harmonic nuclear motions as phonons. Transforming to the interaction representation
and performing a trace over the phonons gives the spectral density in
terms of the autocorrelation of the electron-phonon coupling
operators.
\begin{eqnarray}
S_{nm}(\tilde\omega) = \int_{-\infty}^{\infty} dt e^{-i\tilde \omega t}\langle \hat V_{nm}(t) \hat V_{mn}(0)\rangle.\label{spec-dens}
\end{eqnarray}
Here, $\hat V_{nm}(t)$ is the electron-phonon coupling term in the Heisenberg representation and
$\langle \cdots \rangle$ denotes a thermal average over the
vibrational degrees of freedom.
The derivation and explicit form for the kernel in Eq.~\ref{spec-dens}  is quite lengthy and is given in
Ref.~\cite{pereverzev2006time}.

\subsection{Non-Markovian Master Equation and Golden-Rule Rates}

@@

At this point it is useful to connect the various terms in the phonon-dressed Hamiltonian
with specific physical parameters.
First, the reorganization energy is given by
%The terms involving $$
%are related to the reorganized energy difference between the two states.
%$$
%\lambda_{nm} =\sum_{j}((g_{nnj}-g_{mmj})/\omega_j)^{2} = \sum_{j}\hbar \omega_{j}S_{j}
%$$
\begin{eqnarray}
\lambda_{nm}=\sum_{j}\frac{\left(g_{nnj}-g_{mmj}\right)^{2}}{\omega_{j}} = \sum_{j}\hbar \omega_{j}S_{j}
\end{eqnarray}
where  the $\{S_{j}\}$  are  the Huang-Rhys factors for each phonon mode.
These are related to the Franck-Condon factor
describing the overlap between the $v_j=1$ vibronic state in one electronic state
with the $v_j=0$ vibronic state in the other.
Likewise, the energy difference between the renormalized energy gaps is related to the
driving force of the state-to-state transition,
\begin{eqnarray}
\Delta E_{nm} = \tilde \epsilon_n-\tilde \epsilon_m.
\end{eqnarray}

We first developed the approach about 10 years ago. However, the difficulty in using this approach is that
it requires both diagonal (${\mathbf g}_{nn}$) and off-diagonal  (${\mathbf g}_{nm}$) derivative couplings between
adiabatic states.    In the following section, we discuss how we have used the Edmiston-Ruedenberg (ER) localization scheme \cite{edmiston1963localized} to estimate the couplings.   We also present how one can construct a reduced set of harmonic modes that
fully capture the electron/nuclear coupling.

\subsection{Adiabatic and Diabatic Representations}
There are two common ways to represent the Hamiltonian, adiabatic and diabatic representations, as sketched in Fig. \ref{adiaDiab}. The way we represent the Hamiltonian in Eq. \ref{ham1} is adiabatic representation, where states are the eigenstates of molecule, and hence can be directly calculated in {\em ab initio} quantum chemistry software. However, the off-diagonal couplings in adiabatic representations ($g_{ij}$ in Eq. \ref{ham1}) are derivative couplings, which are difficult to compute. Even worse, the  couplings change radically near avoided crossing. A workaround is switching to diabatic representation, where the Hamiltonian is written as
\begin{eqnarray}
H_{dia}=U^{T}H_{adia}U=\left(\begin{array}{cc}
\epsilon_{a}(\mathbf{R})\mathrm{+T'_{n}}(\mathbf{R})_{11} & V_{ab}\\
V_{ab} & \epsilon_{b}(\mathbf{R})\mathrm{+T'_{n}}(\mathbf{R})_{22}
\end{array}\right).
\end{eqnarray}

The diabatic representation has several advantages. First, sharp derivative couplings in adiabatic representations are transformed to smoother static diabatic couplings, $V_{ab}$. Second, potential energy surfaces are  smoothened and avoided crossing is eliminated.

\begin{figure}[!t]
\includegraphics[width=0.95\columnwidth]{Chapters/chap2/adia_vs_diabatic.png}
\caption{Sketch of adiabatic and diabatic representations for two-state system. Compared to adiabatic representation, diabatic representation has smoother energy surfaces and couplings.
}
\label{adiaDiab}
\end{figure}

The problem now is how to obtain the transformation matrix
\begin{eqnarray}
U = \left(\begin{array}{cc}
\cos\theta & \sin\theta\\
-\sin\theta & \cos\theta
\end{array}\right).
\end{eqnarray}
There are many methods available \cite{domcke2004conical}. Here we introduce the Edmiston-Ruedenberg diabatization approach.

\subsection{Edmiston-Ruedenberg Diabatization}\label{sec:ER}
A straightforward way to construct diabatic states is to eliminate derivative coupling mathematically, namely, to solve
\begin{eqnarray}
\left\langle \phi_{i}(\mathbf{r};\mathbf{R})|\nabla_{\mathbf{R}}|\phi_{j}(\mathbf{r};\mathbf{R})\right\rangle = 0.
\end{eqnarray}
However, it is computationally very expensive, and exact solution do not usually exist \cite{mead1982conditions}.

Another category of diabatization methods is based on physical intuition, with ER diabatization being one of them. The ER diabatization is based on the idea that good diabatic states are supposed to be localized, so the approach defines the diabatic states by minimizing the total electron repulsion
\begin{eqnarray}
f_{ER}=\sum_{k}^{N_{states}}\iint dr_{1}dr_{2}\frac{\left\langle \phi_{k}|\hat{\rho}\left(r_{1}\right)|\phi_{k}\right\rangle \left\langle \phi_{k}|\hat{\rho}\left(r_{2}\right)|\phi_{k}\right\rangle }{\left\Vert r_{1}-r_{2}\right\Vert }.
\end{eqnarray}

\subsection{Parameterization From Quantum Chemistry}

@@

\subsection{Lanczos Method}
It is crucial to notice that the vectors  given in Eq.~\ref{eq:locaiHam} are {\em not linearly independent}.
 Consequently, special care
must be taken to generate the reduced sub-space.  To facilitate this, we develop an iterative Lanczos approach,
taking the normalized vector ${\mathbf v}_{1} = {\mathbf g}_{22}$ as a starting point.

As above, we initialize each step indexed by $k$,  by defining a projection operator
\begin{eqnarray}
{\mathbf P}_{k} = {\mathbf v}_{k}\otimes  {\mathbf v}_{k}
\end{eqnarray}
and its complement ${\mathbf Q}_{k} = {\mathbf I}  - {\mathbf P}_{k}$.
%${\mathbf P}_{k}$ is the projection operator for the
$k$-th mode.
%We also construct
%\begin{eqnarray}
%\mathbf p = \sum_{k} \mathbf P_{k}
%\end{eqnarray}
%as  the total projection operator for all ${k} \le N$ modes.
We then project the Hessian matrix $\mathbf\Omega$ into each subspace {\em viz.}
\begin{eqnarray}
\mathbf\Omega_{p} = \mathbf P_{k}\cdot \mathbf\Omega \cdot \mathbf P_{k} \,\, \& \,\, \mathbf\Omega_{q} = \mathbf Q_{k}\cdot \mathbf\Omega \cdot  \mathbf Q_{k}
\end{eqnarray}
and diagonalize each to obtain eigenvalues and eigenvectors $\{\alpha_{p}, {\mathbf M}_{p}\}$ and $\{\alpha_{q}, {\mathbf M}_{q}\}$
respectively.
As above, $\mathbf\Omega_{p} $ and $\mathbf\Omega_{q}$ are $N\times N$ matrices.
The first set will have a single
non-trivial eigenvalue and the second set
will have $N-k$ non-trivial eigenvalues.  As above we collect the non-trivial eigenvectors associated with each
to form the orthogonal transformation matrix
\begin{eqnarray}
{\mathbf M}_{k} = \{{\mathbf M}_{p},{\mathbf M}_{q}\},
\end{eqnarray}
and again transform the full Hessian $\mathbf\Omega$ into this new vector space to form the $N\times N$ matrix $\mathbf\Omega'$.
 At each step in the iteration, the transformed Hessian, $\mathbf\Omega'$ is in the form of a
$k\times k$ tri-diagonal submatrix in the upper-left part of the matrix and
a diagonal submatrix in the lower-right.  For example, after $k=3$ iterations, the Hessian matrix takes the form:
\begin{eqnarray}
{\mathbf\Omega}' =
\begin{pmatrix}
\alpha_{1}   & b_{1}    & 0                    &               &             &            & 0 \\
b_{1}     & \alpha_{2}  & b_{2}                         \\
0            &   b_{2}         & \alpha_{3}    &   c_{k+1}   & c_{k+2}  & \cdots & c_{N}     \\
               &                     &           c_{k+1}           &     \alpha_{k+ 1} & &             & 0                 \\
               &                     &           c_{k+2}           &                 & \alpha_{k+2}  \\
               &                     &          \vdots           &                 &                    &  \ddots \\
   0           &                     &           c_{N}           &     0   &&& \alpha_{N}\\
\end{pmatrix}.
\label{omega-prime}
\end{eqnarray}
We note that only the $k$-th mode is coupled the $N-k$ remaining modes.
Since all of the transformations are orthogonal, diagonalizing $\mathbf\Omega'$ at any point
returns the original Hessian matrix.

@@

At any point along the way, we can terminate the iteration and obtain a reduced set of
couplings.  Since the  Lanczos approach uses the power method for finding the largest eigenvector of a matrix,
it converges first upon the vector with the largest electron/nuclear coupling--which
 we refer to as the ``primary mode''.  Subsequent iterations produce reduced modes with progressively
 weaker electron/nuclear couplings and the entire  process can be terminated after a few iterations.
After $k$-steps, the final electron-phonon couplings are then obtained by
projecting the original set of couplings (in the normal mode basis) into the final vector space.

For the first iteration, ${\mathbf v}_{1}$ is parallel to the bare electron-phonon coupling vector $g_{22}$
and the associated frequency is ${\mathbf v}_{1}\cdot\Omega\cdot{\mathbf v}_{1}$.   The subsequent iterations introduce
corrections to this via phonon-phonon coupling mediated via the electronic couplings.  For example, for the $k=3$ iteration,
we would determine the active vector space in terms of the upper-left 3$\times 3$ block of the matrix in
Eq.~\ref{omega-prime}.
\begin{eqnarray}
{\mathbf \Omega}_{3}' =
\begin{pmatrix}
\alpha_{1}   & b_{1}    & 0      \\
b_{1}     & \alpha_{2}  & b_{2}                         \\
0            &   b_{2}         & \alpha_{3}
\end{pmatrix}
\end{eqnarray}
Diagonalizing  ${\mathbf \Omega}_{3}'$ returns a set of frequencies and  associated eigenvectors
which are then used to compute the electron-phonon couplings in this reduced active space.
After $N-1$ iterations, $\mathbf\Omega'$ is a fully tridiagonal matrix and diagonalizing this returns the original
normal mode basis.
